### 2022-12-05 DataExploratoryAnalyses.Rmd
Added specific questions to the exploratory analyses

---
title: "Data Exploratory Analyses/Plots"
author: "Kathryn Wheeler"
date: "11/30/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE,results='hide',message=FALSE,warning=FALSE}
source('sharedVariables.R')
source('NEON_Data_DownloadAndProcess.R')
```

```{r,echo=FALSE,results='hide',message=FALSE,warning=FALSE}
library(MuMIn)
library(ggplot2)
library(mgcv)
library(ggridges)

allPrecipData <- read.csv(paste0(dataPath,"NEON_PrecipitationDataDailydata_gapFilled.csv"))
allPrecipData$year <- lubridate::year(allPrecipData$date)
allPrecipData <- allPrecipData %>% group_by(year,siteID) %>% mutate(cumPrecip=cumsum(precipBulk)) #Calculates the cummulative sum of precip for each site-year for each date

allTempData <- read.csv(paste0(dataPath,"NEON_TemperatureDataDailydata_gapFilled.csv"))
allTempData$year <- lubridate::year(allTempData$date)

allTempData$GDDoffset_0 <- ifelse((allTempData$tempTripleMean-0)<0,0,allTempData$tempTripleMean-0)
allTempData <- allTempData %>% group_by(year,siteID) %>% mutate(GDD_0=cumsum(GDDoffset_0))

allTempData$GDDoffset_5 <- ifelse((allTempData$tempTripleMean-5)<0,0,allTempData$tempTripleMean-5)
allTempData <- allTempData %>% group_by(year,siteID) %>% mutate(GDD_5=cumsum(GDDoffset_5))

allTempData$GDDoffset_10 <- ifelse((allTempData$tempTripleMean-10)<0,0,allTempData$tempTripleMean-10)
allTempData <- allTempData %>% group_by(year,siteID) %>% mutate(GDD_10=cumsum(GDDoffset_10))

allTempData$GDDoffset_15 <- ifelse((allTempData$tempTripleMean-15)<0,0,allTempData$tempTripleMean-15)
allTempData <- allTempData %>% group_by(year,siteID) %>% mutate(GDD_15=cumsum(GDDoffset_15))

allTempData <- allTempData[,-c(5,7,9,11)] #Remove the offset columns to just keep cummulative

allSunlightTimes <- read.csv(paste0(dataPath,'NEON_sunlightTimes.csv'))[-c(3,4,5,6)]
```

#For each site-PFT combination: How many individuals of the different associations are present?

```{r,echo=FALSE,results='hide',message=FALSE,warning=FALSE,fig.width=10,fig.height=15}
#Number of Individuals of AM and EcM that are present
uniqueIndividuals <- matrix(nrow=0,ncol=5)
for(p in seq_along(NEON_phenophase_names)){
  #Bud burst 
  print(NEON_phenophase_names[p])

  subDat <- read.csv(file=paste0(dataPath,'NEON_PhenologyObservations/NEON_PhenoObservationData_',gsub(" ","",NEON_phenophase_names[p]),'.csv'))
  subDat <- subDat[,c('siteID','individualID','growthForm','Mycorrhizal.type','decimalLatitude')]
  subDat <- subDat %>% unique()
  uniqueIndividuals <- rbind(uniqueIndividuals,subDat)
  uniqueIndividuals <- uniqueIndividuals %>% unique()
}
uniqueIndividuals <- uniqueIndividuals[!(uniqueIndividuals$siteID%in%c("2019","2020","2021","2017","2018","2022")),]
#uniqueIndividuals <- uniqueIndividuals %>% filter(Mycorrhizal.type %in% c("AM","EcM"))
uniqueIndividuals <- uniqueIndividuals %>% filter(!is.na(growthForm))
uniqueIndividuals[uniqueIndividuals$growthForm=="Pine",'growthForm'] <- "Evergreen conifer"
uniqueIndividuals <- uniqueIndividuals %>% filter(growthForm %in% c("Deciduous broadleaf","Drought deciduous broadleaf","Evergreen broadleaf","Evergreen conifer","Forb","Graminoid","Semi-evergreen broadleaf")) #Exclude the others

uniqueIndividuals2 <- uniqueIndividuals %>% group_by(siteID,Mycorrhizal.type,growthForm,decimalLatitude) %>% summarise(n=n())

ggplot(uniqueIndividuals2,aes(fill=Mycorrhizal.type,y=n,x=fct_reorder(siteID,decimalLatitude))) + 
  geom_bar(position="dodge",stat="identity")+
  facet_wrap(~growthForm,ncol=4)+
  theme_classic()+
  theme(axis.text.x = element_text(angle=90))+
  xlab("Site ID (Ordered by Latitude)")+
  ylab("Number of Individuals")+
  coord_flip()

```

What I learned is that DB has the best variation between different mycorrhizal associations so that's where I should focus my comparitive studies and then do across PFTs for the overall.

#For each site-phenophase combination: how many years are individuals re-measured?
```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.width=8,fig.height=10}
for(p in seq_along(NEON_phenophase_names)){
  subDat <- read.csv(file=paste0(dataPath,'NEON_PhenologyObservations/NEON_PhenoObservationData_',gsub(" ","",NEON_phenophase_names[p]),'.csv'))
  subDat <- subDat[,c('siteID','individualID','Mycorrhizal.type','decimalLatitude','year')]
  subDat <- subDat %>% unique()

  subDat2 <- subDat %>% group_by(siteID,Mycorrhizal.type,individualID,decimalLatitude) %>% summarise(n=n())
  if(nrow(subDat2)>2){
    print(ggplot(subDat2,aes(n,fct_reorder(siteID,decimalLatitude),fill=Mycorrhizal.type)) +
            geom_density_ridges2(scale=1)+
            theme_classic()+
            theme(axis.text.x = element_text(angle=90))+
            xlab("Number Of Years Per Individual")+
            ylab("Site ID (Ordered by Latitude)")+
            ggtitle(NEON_phenophase_names[p])+
            scale_x_continuous(limits=c(0,8)))
  }
}

```

#For each PFT-phenophase combination: Does the Standard Deviation in DOYs and lengths between phenophases differ between EcM and AM associated plants acrross all sites and within each site?

```{r,echo=FALSE,results='hide',message=FALSE,warning=FALSE,fig.width=10,fig.height=20}
for(p in seq_along(NEON_phenophase_names)){
  subDat <- read.csv(file=paste0(dataPath,'NEON_PhenologyObservations/NEON_PhenoObservationData_',gsub(" ","",NEON_phenophase_names[p]),'.csv'))
  subDat <- subDat %>% filter(phenophaseIntensity == mediumIntensity_phenophases[p])
  subDat <- subDat[,c('siteID','individualID','Mycorrhizal.type','decimalLatitude','dayOfYear','growthForm')]
  subDat <- subDat %>% filter(growthForm %in% c("Deciduous broadleaf","Drought deciduous broadleaf","Forb","Graminoid"))
  subDat2 <- subDat %>% group_by(individualID) %>% mutate(doy_sd=sd(dayOfYear,na.rm=TRUE))

  if(nrow(subDat2)>2){
    print(ggplot(subDat2,aes(doy_sd,fct_reorder(siteID,decimalLatitude),fill=Mycorrhizal.type)) +
            geom_density_ridges2(scale=1)+
            theme_bw()+
            theme(axis.text.x = element_text(angle=90))+
            xlab("SD in DOY Per Individual")+
            ylab("Site ID (Ordered by Latitude)")+
            facet_wrap(~growthForm)+
            ggtitle(NEON_phenophase_names[p])+
            scale_x_continuous(limits=c(0,100)))
  }
}

#Lengths Between Phenophases:

```

#For each PFT-phenophase combination: Does the DOY of the timing and the lengths between phenophases differ between EcM and AM assocated plants across all sites and within each site?

```{r}

```

#For each individual and phenophase, what predictors seem to be the most important?

```{r}

```

#Do the basic GAMs work for any of the phenophase and PFTs?

```{r}

```

```{r,eval=FALSE}
p <- 2 #Bud burst 
  print(NEON_phenophase_names[p])
  subDat <- read.csv(file=paste0(dataPath,'NEON_PhenologyObservations/NEON_PhenoObservationData_',gsub(" ","",NEON_phenophase_names[p]),'.csv'))
  subDat <- subDat %>% filter(phenophaseIntensity == mediumIntensity_phenophases[p])
  subDat$year <- as.numeric(subDat$year)
  subDat <- left_join(subDat,allPrecipData[,-3],by=c('siteID','date','year'))
  subDat <- left_join(subDat,allTempData,by=c('siteID','date','year'))
  subDat <- left_join(subDat,allSunlightTimes,by=c('siteID','date'))
  subDat <- subDat %>% filter(year!=2022)
  subDat$Mycorrhizal.type <- as.factor(subDat$Mycorrhizal.type)

  #Deciduous Broadleaf
  subDatDB <- subDat %>% filter(growthForm=="Deciduous broadleaf")
  subDatDB_ECM <- subDatDB %>% filter(Mycorrhizal.type=="EcM",growthForm=="Deciduous broadleaf")
  subDatDB_AM <- subDatDB %>% filter(Mycorrhizal.type=="AM",growthForm=="Deciduous broadleaf")

  plot(density(subDatDB_AM$dayOfYear),main="Day of Year of Budburst for DB")
  lines(density(subDatDB_ECM$dayOfYear),col="red")
  legend('topleft',col=c("black","red"),legend = c("AM DB","ECM DB"),lty=1)

  ggplot(subDatDB,aes(as.numeric(as.character(dayOfYear)),fct_reorder(siteID,decimalLatitude),fill=Mycorrhizal.type)) +
  geom_density_ridges2(scale=1)+
  theme_classic()+
  theme(axis.text.x = element_text(angle=90))+
  xlab("Day of Year")+
  ylab("Site ID (Ordered by Latitude)")+
    ggtitle("Day of Year of Bud Burst in Deciduous Broadleaf")+
  scale_x_continuous(limits=c(0,200))

    ggplot(subDatDB,aes(as.numeric(as.character(dayOfYear)),Genus,fill=Mycorrhizal.type)) +
  geom_density_ridges2(scale=1)+
  theme_classic()+
  theme(axis.text.x = element_text(angle=90))+
  xlab("Day of Year")+
  ylab("Site ID (Ordered by Latitude)")+
    ggtitle("Day of Year of Bud Burst in Deciduous Broadleaf")+
  scale_x_continuous(limits=c(0,200))

    output <- matrix(ncol=length(unique(subDatDB$Genus)),nrow=length_NEON_siteNames)
    for(s in seq_along(NEON_siteNames)){
      print(NEON_siteNames[s])
      for(j in seq_along(unique(subDatDB$Genus))){
        output[s,j] <- nrow(subDatDB %>% 
                              filter(Genus==(unique(subDatDB$Genus)[j]),siteID==NEON_siteNames[s]))
      }
    }
    speciesPerSite <- data.frame(output)
    colnames(speciesPerSite) <- unique(subDatDB$Genus)
    speciesPerSite <- cbind(NEON_siteNames,speciesPerSite)

    MTperSite <- data.frame(matrix(ncol=6,nrow=length_NEON_siteNames))
    colnames(MTperSite) <- c("siteID","Total_AM_ind","Total_ECM_ind","Total_AM_gen","Total_ECM_gen","Lat")
    MTperSite$siteID <- NEON_siteNames
    for(s in seq_along(NEON_siteNames)){
      print(NEON_siteNames[s])
      siteDat <- subDatDB %>% filter(siteID==NEON_siteNames[s])
      MTperSite$Total_AM_ind[s] <- sum(siteDat$Mycorrhizal.type=="AM")
      MTperSite$Total_ECM_ind[s] <- sum(siteDat$Mycorrhizal.type=="EcM")
      MTperSite$Total_AM_gen[s] <- length(unique((siteDat %>% filter(Mycorrhizal.type=="AM"))$Genus))
      MTperSite$Total_ECM_gen[s] <- length(unique((siteDat %>% filter(Mycorrhizal.type=="EcM"))$Genus))
      MTperSite$Lat[s] <- as.numeric(siteData[siteData$field_site_id==NEON_siteNames[s],'field_latitude'])
    }
    MTperSite <- MTperSite[order(MTperSite$Lat,decreasing = TRUE),]

    mdl_gam <- gam(dayOfYear ~ s(decimalLatitude,k=25) + Mycorrhizal.type,data=subDatDB,
                   method="REML",
                   na.action="na.fail")
    plot.gam(mdl_gam,residuals=TRUE,pch=20,cex=5,shade=TRUE,page=1,all.terms = TRUE)
      gam.check(mdl_gam,page=1)
    plot(subDatDB$decimalLatitude,subDatDB$dayOfYear,pch=20,xlab="Latitude",ylab="Day Of Year of Budburst")
    abline(lm(dayOfYear ~ decimalLatitude,data=subDatDB),lwd=3)
    abline(lm(dayOfYear ~ decimalLatitude,data=subDatDB[subDatDB$Mycorrhizal.type=="AM",]),lwd=3,col="cyan")
    abline(lm(dayOfYear ~ decimalLatitude,data=subDatDB[subDatDB$Mycorrhizal.type=="EcM",]),lwd=3,col="red")
    legend('bottomright',col=c("black","cyan","red"),legend = c("All DB","AM DB","ECM DB"),lty=1,lwd=3)

  concurvity(mdl_gam,full=TRUE)

  ggplot(subDatDB,aes(as.numeric(as.character(GDD_10)),fct_reorder(siteID,decimalLatitude),fill=Mycorrhizal.type)) +
  geom_density_ridges2(scale=1)+
  theme_classic()+
  theme(axis.text.x = element_text(angle=90))+
  xlab("GDD_10 (C)")+
  ylab("Site ID (Ordered by Latitude)")+
    ggtitle("GDD_10 of Bud Burst in Deciduous Broadleaf")+
  scale_x_continuous(limits=c(0,600))

      ggplot(subDatDB,aes(as.numeric(as.character(GDD_10)),as.numeric(as.character(dayOfYear)),col=Mycorrhizal.type))+geom_point(cex=.1)+
        facet_wrap(~siteID,ncol=6)+
  theme_classic()+
  theme(axis.text.x = element_text(angle=90))+
  xlab("GDD_10 (C)")+
  ylab("Day of Year of Budburst")+
        geom_smooth(method="lm")+ 
  scale_x_continuous(limits=c(0,1500))

  gdd_doy_slopes <- data.frame(matrix(nrow=length_NEON_siteNames,ncol=3))
  colnames(gdd_doy_slopes) <- c("siteID","AM_slope","ECM_slope")
  gdd_doy_slopes$siteID <- NEON_siteNames
  for(s in seq_along(NEON_siteNames)){
      print(NEON_siteNames[s])
      siteDat <- subDatDB %>% filter(siteID==NEON_siteNames[s])
      if(sum(siteDat$Mycorrhizal.type=="AM")>1){
      gdd_doy_slopes$AM_slope[s] <- summary(lm(dayOfYear~GDD_10,
                                               data=siteDat[siteDat$Mycorrhizal.type=="AM",]))$coefficients[2,1]
      }
      if(sum(siteDat$Mycorrhizal.type=="EcM")>3){
      gdd_doy_slopes$ECM_slope[s] <- summary(lm(dayOfYear~GDD_10,
                                               data=siteDat[siteDat$Mycorrhizal.type=="EcM",]))$coefficients[2,1]
      }
  }
  plot(density(na.omit(gdd_doy_slopes$AM_slope)),col="cyan",main="Slopes of Site-Specific Linear Regressions Between GDD_10 and DOY of Bud Burst")
  lines(density(na.omit(gdd_doy_slopes$ECM_slope)),col="red")
  legend('topright',col=c("cyan","red"),legend = c("AM DB","ECM DB"),lty=1,lwd=3)

  dl_doy_slopes <- data.frame(matrix(nrow=length_NEON_siteNames,ncol=3))
  colnames(dl_doy_slopes) <- c("siteID","AM_slope","ECM_slope")
  dl_doy_slopes$siteID <- NEON_siteNames
  for(s in seq_along(NEON_siteNames)){
      print(NEON_siteNames[s])
      siteDat <- subDatDB %>% filter(siteID==NEON_siteNames[s])
      if(sum(siteDat$Mycorrhizal.type=="AM")>1){
      dl_doy_slopes$AM_slope[s] <- summary(lm(dayOfYear~daylength,
                                               data=siteDat[siteDat$Mycorrhizal.type=="AM",]))$coefficients[2,1]
      }
      if(sum(siteDat$Mycorrhizal.type=="EcM")>3){
      dl_doy_slopes$ECM_slope[s] <- summary(lm(dayOfYear~daylength,
                                               data=siteDat[siteDat$Mycorrhizal.type=="EcM",]))$coefficients[2,1]
      }
  }
  plot(density(na.omit(dl_doy_slopes$AM_slope)),col="cyan",main="Slopes of Site-Specific Linear Regressions Between Day length and DOY of Bud Burst")
  lines(density(na.omit(dl_doy_slopes$ECM_slope)),col="red")
  legend('topright',col=c("cyan","red"),legend = c("AM DB","ECM DB"),lty=1,lwd=3)

  mdl_gam <- gam(dayOfYear ~ s(daylength, by=as.factor(siteID)) + Mycorrhizal.type,data=subDatDB,
                 method="REML",
                 na.action="na.fail")
    plot.gam(mdl_gam,residuals=TRUE,pch=20,cex=5,shade=TRUE,page=1,all.terms = TRUE)
    gam.check(mdl_gam)

    plot(subDatDB$daylength,subDatDB$dayOfYear,pch=20)
    points(subDatDB[subDatDB$Mycorrhizal.type=="EcM",'daylength'],
           subDatDB[subDatDB$Mycorrhizal.type=="EcM",'dayOfYear'],pch=20,col="red")

    points(subDatDB[subDatDB$Mycorrhizal.type=="AM",'daylength'],
           subDatDB[subDatDB$Mycorrhizal.type=="AM",'dayOfYear'],pch=20,col="cyan")

    subDatDB2 <- subDatDB %>% group_by(individualID) %>% mutate(doy_sd=sd(dayOfYear,na.rm=TRUE))
    subDatDB2 <- subDatDB2[,c('siteID','individualID','doy_sd','decimalLatitude','Mycorrhizal.type')] %>% unique()

    ggplot(subDatDB2,aes(as.numeric(as.character(doy_sd)),
                         fct_reorder(siteID,decimalLatitude),fill=Mycorrhizal.type)) +
  geom_density_ridges2(scale=1)+
  theme_classic()+
  theme(axis.text.x = element_text(angle=90))+
  xlab("Day of Year SD")+
  ylab("Site ID (Ordered by Latitude)")+
    ggtitle("SD in Day Of Year of Budburst in Deciduous Broadleaf")+
  scale_x_continuous(limits=c(0,30))

    plot(density(na.omit(subDatDB2$doy_sd)),main="SD in DOY of bud burst for DB individuals",xlim=c(0,40))
    lines(density(na.omit(as.numeric(subDatDB2[subDatDB2$Mycorrhizal.type=="AM",]$doy_sd))),col="cyan")
    lines(density(na.omit(as.numeric(subDatDB2[subDatDB2$Mycorrhizal.type=="EcM",]$doy_sd))),col="red")
    legend('topright',col=c("black","cyan","red"),legend = c("All DB","AM DB","ECM DB"),lty=1,lwd=3)

```

```{r,eval=FALSE}
p <- 3 #Senescence
  print(NEON_phenophase_names[p])
  subDat <- read.csv(file=paste0(dataPath,'NEON_PhenologyObservations/NEON_PhenoObservationData_',gsub(" ","",NEON_phenophase_names[p]),'.csv'))
  subDat <- subDat %>% filter(phenophaseIntensity == mediumIntensity_phenophases[p])
  subDat$year <- as.numeric(subDat$year)
  subDat <- left_join(subDat,allPrecipData[,-3],by=c('siteID','date','year'))
  subDat <- left_join(subDat,allTempData,by=c('siteID','date','year'))
  subDat <- left_join(subDat,allSunlightTimes,by=c('siteID','date'))
  subDat <- subDat %>% filter(year!=2022)
  subDat$Mycorrhizal.type <- as.factor(subDat$Mycorrhizal.type)

  #Deciduous Broadleaf
  subDatDB <- subDat %>% filter(growthForm=="Deciduous broadleaf")
  subDatDB_ECM <- subDatDB %>% filter(Mycorrhizal.type=="EcM",growthForm=="Deciduous broadleaf")
  subDatDB_AM <- subDatDB %>% filter(Mycorrhizal.type=="AM",growthForm=="Deciduous broadleaf")

    ggplot(subDatDB,aes(as.numeric(as.character(dayOfYear)),fct_reorder(siteID,decimalLatitude),fill=Mycorrhizal.type)) +
  geom_density_ridges2(scale=1)+
  theme_classic()+
  theme(axis.text.x = element_text(angle=90))+
  xlab("Day of Year")+
  ylab("Site ID (Ordered by Latitude)")+
    ggtitle("Day of Year of Senescence in Deciduous Broadleaf")+
  scale_x_continuous(limits=c(150,400))

  ggplot(subDatDB,aes(as.numeric(as.character(GDD_15)),fct_reorder(siteID,decimalLatitude),fill=Mycorrhizal.type)) +
  geom_density_ridges2(scale=1)+
  theme_classic()+
  theme(axis.text.x = element_text(angle=90))+
  xlab("GDD_15 (C)")+
  ylab("Site ID (Ordered by Latitude)")+
    ggtitle("GDD_10 of Senescence in Deciduous Broadleaf")

```