### 2023-01-25 manuscriptCode_phenologyAndMycorrhizae.Rmd

---
title: "Phenology and Mycorrhizae Manuscript"
author: "Kathryn Wheeler"
date: "11/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load Functions and Variables
```{r}
source('sharedVariables.R')
source('NEON_Data_DownloadAndProcess.R')
```

Download NEON Data

#NEON Phenology Observations 
```{r,eval=FALSE}
source('combineNEONdata_PhenoObs.R')
dataName="NEON_PhenologyObservations"
NEON_ID='DP1.10055.001'

downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID)
combineNEONdata_PhenoObs(saveType=="bySite") #Saved in individual files by site because 1 file was too large 
combineNEONdata_PhenoObs(saveType=="firstOfPhenophase")
combineNEONdata_PhenoObs(saveType = "allOfPhenophase")
```

#Single Air Temperature At Various Heights 
```{r,eval=FALSE}
dataName="NEON_SingleAirTemperature"
NEON_ID='DP1.00002.001'
selectColumns <- c('horizontalPosition','verticalPosition','tempSingleMean','tempSingleExpUncert','finalQF')
inFileName <- "SAAT_30min.csv"

downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID)
combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,selectColumns=selectColumns,
                inFileName=inFileName,dataPath=dataPath)

```

#Triple Air Temperature at Top of Canopy
```{r,eval=FALSE} 
##Need to redownload!! 
dataName="NEON_TemperatureData"
NEON_ID='DP1.00003.001'
selectColumns <- c('tempTripleMean','finalQF')
inFileName <- "TAAT_30min.csv"
varName <- 'tempTripleMean'

downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID)
combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,selectColumns=selectColumns,
                inFileName=inFileName,dataPath=dataPath)
calculateDailyWeather(dataName=dataName,dataPath=dataPath,varName=varName,funType=mean)
gapFillFromDaymet(dataName=dataName,dataPath=dataPath,varName=varName)

```

#Precipitation
```{r,eval=FALSE}
dataName="NEON_PrecipitationData"
NEON_ID='DP1.00006.001'
selectColumns <- c('precipBulk','precipExpUncert','precipQF')
inFileName <- c("PRIPRE_30min.csv","SECPRE_30min.csv")
varName <- "precipBulk"

downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID)
combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,selectColumns=selectColumns,
                inFileName=inFileName,dataPath=dataPath)
calculateDailyWeather(dataName=dataName,dataPath=dataPath,varName=varName,funType=sum)
gapFillFromDaymet(dataName=dataName,dataPath=dataPath,varName=varName)

```

#Download Photosynthetically active radiation (PAR)
```{r}
dataName="NEON_PAR"
NEON_ID='DP1.00024.001'
if(!file.exists(paste0(dataPath,"/",dataName))){
  dir.create(paste0(dataPath,"/",dataName))
}
selectColumns <- c('verticalPosition','PARMean','PARMinimum','PARMaximum','PARVariance','PARFinalQF')
inFileName <- "PARPAR_30min.csv"
# varName <- "precipBulk"

downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID)
combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,selectColumns=selectColumns,
                inFileName=inFileName,dataPath=dataPath)
calculateDailyWeather(dataName=dataName,dataPath=dataPath,varName=varName,funType=sum)

```

#Download NEON Soil Microbe Data
```{r,eval=FALSE}
dataName="NEON_SoilMicrobeComposition"
NEON_ID='DP1.10081.001'
#selectColumns <-
inFileName <- c("mcc_soilSeqVariantMetadata_16S.csv","mcc_soilSeqVariantMetadata_ITS.csv")
#varName <-

downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID) #Note: I had to modify this function to remove the indices for sites that didn't seem to have data. 
combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,selectColumns=selectColumns,
                inFileName=inFileName,dataPath=dataPath)
```

#Download NEON Soil Physical and Chemical Properties (Need to finish combining)
```{r,eval=FALSE}
dataName <- "NEON_soilProperties"
NEON_ID <- "DP1.10086.001"
if(!file.exists(paste0(dataPath,"/",dataName))){
  dir.create(paste0(dataPath,"/",dataName))
}

downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID)

inFileName <- "sls_soilCoreCollection.csv"
selectColumns <- c('plotID','decimalLatitude','decimalLongitude','elevation','standingWaterDepth','horizon',
                   'soilTemp','litterDepth')
dat1 <- combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,selectColumns=selectColumns,
                inFileName=inFileName,dataPath=dataPath,saveFile = FALSE)

dat1 <- dat1 %>% mutate(collectDate=format(as.POSIXct(collectDate,                                                  format='%Y-%m-%dT%H:%M'),format="%Y-%m-%d")) %>%
  group_by(siteID,collectDate,plotID,decimalLatitude,decimalLongitude,elevation,horizon) %>% 
  summarise(soilTemp=mean(soilTemp),litterDepth=mean(litterDepth))

inFileName <- "sls_soilpH.csv"
selectColumns <- c('plotID','horizon','soilInWaterpH','soilInCaClpH')
dat2 <- combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,selectColumns=selectColumns,
                inFileName=inFileName,dataPath=dataPath,saveFile = FALSE)

dat2 <- dat2 %>% mutate(collectDate=format(as.POSIXct(collectDate,                                                  format='%Y-%m-%dT%H:%M'),format="%Y-%m-%d")) %>% group_by(siteID,collectDate,plotID,horizon) %>%
  summarise(soilInWaterpH=mean(soilInWaterpH),soilInWaterpH=mean(soilInWaterpH))

allDat <- full_join(dat1,dat2,by=c('siteID','collectDate','plotID','horizon'))

inFileName <- "sls_soilMoisture.csv"
selectColumns <- c('plotID','horizon','soilMoisture','dryMassFraction')
dat3 <- combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,selectColumns=selectColumns,
                inFileName=inFileName,dataPath=dataPath,saveFile = FALSE)
dat3 <- dat3 %>% mutate(collectDate=format(as.POSIXct(collectDate,                                                  format='%Y-%m-%dT%H:%M'),format="%Y-%m-%d"))%>% group_by(siteID,collectDate,plotID,horizon) %>%
  summarise(soilMoisture=mean(soilMoisture),dryMassFraction=mean(dryMassFraction)) 
allDat <- full_join(allDat,dat3,by=c('siteID','collectDate','plotID','horizon'))

outFileName <- paste0(dataName,"ALLdata.csv")
write.csv(file=paste0(dataPath,outFileName),allDat,row.names = FALSE,quote=FALSE)
```

#Download Soil Temp
```{r}
dataName="NEON_SoilTemp"
NEON_ID='DP1.00041.001'
if(!file.exists(paste0(dataPath,"/",dataName))){
  dir.create(paste0(dataPath,"/",dataName))
}

downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID,andStack=FALSE) #missing 30

selectColumns <- c('horizontalPosition','verticalPosition','soilTempMean','soilTempMinimum','soilTempMaximum','finalQF')
inFileName <- "ST_30_minute.csv"

combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,selectColumns=selectColumns,
                inFileName=inFileName,dataPath=dataPath)

```

#NEON Root Characteristics
```{r, eval=FALSE}
dataName="NEON_Roots"
NEON_ID='DP1.10067.001'

if(!file.exists(paste0(dataPath,"/",dataName))){
  dir.create(paste0(dataPath,"/",dataName))
}

downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID) #missing 30

inFileName <- 'bbc_dilution.csv'
selectColumns <- c('plotID','sampleVolume',
                   'dryMass','somDryMass')

dat1 <- combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,
                       selectColumns=selectColumns, inFileName=inFileName,dataPath=dataPath,saveFile = FALSE)
dat1 <- dat1 %>% mutate(density=dryMass/sampleVolume,
                        somDensity=somDryMass/sampleVolume) %>%
  dplyr::select(-c(sampleVolume,dryMass,somDryMass)) %>%
  group_by(siteID,collectDate,plotID) %>% summarise(densityMean=mean(density),somDensityMean=mean(somDensity))

inFileName <- 'bbc_percore.csv'
selectColumns <- c('plotID','decimalLatitude','decimalLongitude',
                   'litterDepth')

dat2 <- combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,
                       selectColumns=selectColumns, inFileName=inFileName,dataPath=dataPath,saveFile = FALSE)

dat2 <- dat2 %>% group_by(siteID,collectDate,plotID) %>%
  summarise(litterDepth=mean(litterDepth))

allDat <- full_join(dat1,dat2,by=c('siteID','collectDate','plotID'))

inFileName <- 'bbc_rootChemistry.csv'
selectColumns <- c('plotID','plotType',
                   'd15N','d13C','nitrogenPercent','carbonPercent',
                   'CNratio','cnIsotopeQF','cnPercentQF')

dat3 <- combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,
                       selectColumns=selectColumns, inFileName=inFileName,dataPath=dataPath,saveFile = FALSE)

dat3 <- dat3 %>% mutate(collectDate=format(as.POSIXct(collectDate,                                                  format='%Y-%m-%dT%H:%M'),format="%Y-%m-%d")) %>%
  filter(cnIsotopeQF=="OK",cnPercentQF=="OK") %>%
  group_by(siteID,collectDate,plotID) %>% summarise(d15N=mean(d15N),
                                                    d13C=mean(d13C),
                                                    nitrogenPercent=
                                                      mean(nitrogenPercent),
                                                    carbonPercent=
                                                      mean(carbonPercent),
                                                    CNratio=
                                                      mean(CNratio))

allDat <- full_join(allDat,dat3,by=c('siteID','collectDate','plotID'))

inFileName <- "bbc_rootmass.csv"
selectColumns <- c('plotID','dryMass','mycorrhizaeVisible')

dat4 <- combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,
                       selectColumns=selectColumns, inFileName=inFileName,dataPath=dataPath,saveFile = FALSE)
names(dat4)[names(dat4)=="dryMass"] <- "rootDryMass"
dat4$mycorrhizaeVisible[dat4$mycorrhizaeVisible=="N"] <- 0
dat4$mycorrhizaeVisible[dat4$mycorrhizaeVisible=="Y"] <- 1
dat4$mycorrhizaeVisible[dat4$mycorrhizaeVisible==""] <- NA

dat4 <- dat4 %>% group_by(siteID,collectDate,plotID) %>%
  summarise(rootDryMass=mean(rootDryMass),
            mycorrhizaeVisiblePercent=mean(as.numeric(mycorrhizaeVisible),na.rm=TRUE))

allDat <- full_join(allDat,dat4,by=c('siteID','collectDate','plotID'))
outFileName <- paste0(dataName,"ALLdata.csv")
write.csv(file=paste0(dataPath,outFileName),allDat,row.names = FALSE,quote=FALSE)
```

#NEON Plant Structural Characteristics
```{r}

```

#NEON Litterfall
```{r}
dataName="NEON_litterfall" #No: 2, 7,8, 11, 17, 18, 20, 23, 25, 27, 38, 41, 45
NEON_ID="DP1.10033.001"
if(!file.exists(paste0(dataPath,"/",dataName))){
  dir.create(paste0(dataPath,"/",dataName))
}

downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID)

inFileName <- "ltr_litterCarbonNitrogen.csv"

selectColumns <- c('plotID','plotType','setDate','d15N','d13C',
                   'nitrogenPercent','carbonPercent','CNratio')

dat1 <- combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,
                        selectColumns=selectColumns, inFileName=inFileName,dataPath=dataPath,saveFile = FALSE)

dat1 <- dat1 %>% mutate(collectDate=format(as.POSIXct(collectDate,                                                  format='%Y-%m-%dT%H:%M'),format="%Y-%m-%d")) %>%
  group_by(siteID,collectDate,plotID,plotType,setDate) %>%
  summarise(d15N=mean(d15N), d13C=mean(d13C),nitrogenPercent=mean(nitrogenPercent),
            carbonPercent=mean(carbonPercent), CNratio=mean(CNratio))

inFileName <- "ltr_litterLignin.csv"
selectColumns <- c('plotID','setDate','ligninPercent',
                   'cellulosePercent')

dat2 <- combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,
                        selectColumns=selectColumns, inFileName=inFileName,dataPath=dataPath,saveFile = FALSE)

dat2 <- dat2 %>% mutate(collectDate=format(as.POSIXct(collectDate,                                                  format='%Y-%m-%dT%H:%M'),format="%Y-%m-%d")) %>%
  group_by(siteID,collectDate,plotID,setDate) %>%
  summarise(ligninPercent=mean(ligninPercent),
            cellulosePercent=mean(cellulosePercent))

allDat <- full_join(dat1,dat2,by=c('siteID','plotID','setDate'))

inFileName <- "ltr_massdata.csv"
selectColumns <- c('plotID','setDate','functionalGroup',
                   'dryMass')

dat3 <- combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,
                        selectColumns=selectColumns, inFileName=inFileName,dataPath=dataPath,saveFile = FALSE)
dat3 <- dat3 %>%
  group_by(siteID,collectDate,plotID,setDate,functionalGroup) %>%
  summarise(dryMass=mean(dryMass))

allDat <- full_join(allDat,dat3,by=c('siteID','plotID','setDate'))

outFileName <- paste0(dataName,"ALLdata.csv")
write.csv(file=paste0(dataPath,outFileName),allDat,row.names = FALSE,quote=FALSE)

```

#NEON Foliar Traits
```{r}
dataName="NEON_plantFoliarTraits"
NEON_ID="DP1.10026.001"

if(!file.exists(paste0(dataPath,"/",dataName))){
  dir.create(paste0(dataPath,"/",dataName))
}

downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID) #missing 30

inFileName <- "cfc_LMA.csv"
selectColumns <- c('sampleID','plotType','percentGreen','leafMassPerArea','dryMassFraction')

dat1 <- combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,
                        selectColumns=selectColumns, inFileName=inFileName,dataPath=dataPath,saveFile = FALSE)

inFileName <- "cfc_carbonNitrogen.csv"
selectColumns <- c('plotID','sampleID','plotType','d15N','d13C',
                   'nitrogenPercent','carbonPercent','CNratio')

dat2 <- combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,
                        selectColumns=selectColumns, inFileName=inFileName,dataPath=dataPath,saveFile = FALSE)

inFileName <- "cfc_chlorophyll.csv"
# selectColumns <- c('plotID','chlCarotWavelength1'),'chlCarotWavelength2',
#                    'chlCarotWavelength3','chlCarotWavelength4',
#                    'chlCarotWavelength1Abs','chlCarotWavelength2Abs',
#                    'chlCarotWavelength3Abs','chlCarotWavelength4Abs',
#                    'extractChlAConc','extractChlBConc')
selectColumns <- c('plotID','extractChlAConc','extractChlBConc','extractCarotConc','sampleID')

dat3 <- combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,
                        selectColumns=selectColumns, inFileName=inFileName,dataPath=dataPath,saveFile = FALSE)

inFileName <- "cfc_elements.csv"
selectColumns <- c("plotID",'sampleID','foliarPhosphorusConc','foliarPotassiumConc',
                   'foliarMagnesiumConc','foliarSulfurConc',
                   'foliarManganeseConc','foliarIronConc',
                   'foliarCopperConc','foliarBoronConc','foliarZincConc')

dat4 <- combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,
                        selectColumns=selectColumns, inFileName=inFileName,dataPath=dataPath,saveFile = FALSE)

inFileName <- "cfc_lignin.csv"
selectColumns <- c('plotID','sampleID','plotType','ligninPercent','cellulosePercent')

dat5 <- combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,
                        selectColumns=selectColumns, inFileName=inFileName,dataPath=dataPath,saveFile = FALSE)

allDat <- full_join(dat1,dat2,by=c('siteID','collectDate','sampleID','plotType'))
allDat <- full_join(allDat,dat3,by=c('siteID','collectDate','sampleID','plotID'))
allDat <- full_join(allDat,dat4,by=c('siteID','collectDate','sampleID','plotID'))
allDat <- full_join(allDat,dat5,by=c('siteID','collectDate','sampleID','plotID','plotType'))
##
allDat <- full_join(dat2,dat3,by=c('siteID','plotID','collectDate','sampleID'))
allDat <- full_join(allDat,dat4,by=c('siteID','plotID','collectDate','sampleID'))
allDat <- full_join(allDat,dat5,by=c('siteID','plotID','collectDate','sampleID','plotType'))
allDat <- full_join(allDat,dat1,by=c('siteID','collectDate','sampleID','plotType'))

allDat <- allDat %>% dplyr::select(-sampleID) %>%
  mutate(collectDate=format(as.POSIXct(collectDate,                                                  format='%Y-%m-%dT%H:%M'),format="%Y-%m-%d")) %>%
  group_by(siteID,collectDate,plotID,plotType) %>%
  summarise(percentGreen=mean(percentGreen),leafMassPerArea=mean(leafMassPerArea),
            dryMassFraction=mean(dryMassFraction),
            d15N=mean(d15N),
            d13C=mean(d13C),
            nitrogenPercent=mean(nitrogenPercent),
            carbonPercent=mean(carbonPercent),
            CNratio=mean(CNratio),
            extractChlAConc=mean(extractChlAConc),
            extractChlBConc=mean(extractChlBConc),
            extractCarotConc=mean(extractCarotConc),
            foliarPhosphorusConc=mean(foliarPhosphorusConc),
            foliarPotassiumConc=mean(foliarPotassiumConc),
            foliarMagnesiumConc=mean(foliarMagnesiumConc),
            foliarSulfurConc=mean(foliarSulfurConc),
            foliarManganeseConc=mean(foliarManganeseConc),
            foliarIronConc=mean(foliarIronConc),
            foliarCopperConc=mean(foliarCopperConc),
            foliarBoronConc=mean(foliarBoronConc),
            foliarZincConc=mean(foliarZincConc),
            ligninPercent=mean(ligninPercent),
            cellulosePercent=cellulosePercent)

outFileName <- paste0(dataName,"ALLdata.csv")
write.csv(file=paste0(dataPath,outFileName),allDat,row.names = FALSE,quote=FALSE)

```

#NEON PhenoCam Observations
```{r,eval=FALSE}
source('determinePhenoCamURLs.R')
source('downloadPhenoCamAndCalculateTransitions.R') #Still need to change to have it save it somewhere

```